---
title: "Aides Drug"
author: " Sarh Castellon"
date: "October 8 2024"
output: html_document
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
triUmp <-read.csv(file="triUmpsal.csv",header=TRUE)
```

# Introduction

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?.  If you skimp in this section expect to be marked down in your grade. Paint a picture of the project a couple of paragraphs is expected!-->

A new  preventative aides drug (triUmpsal) is under test in clinical trials. Consider only the females, does this drug seem effective?  Hint - use dplyr to obtain that subpopulation with the filter command.

# Methods

$H_0$ The vaccine is not effective in preventing women from contracting aids  
$H_A$ The vaccine is effective in preventing women from contracting aids  

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method. Discuss what data manipulations may be needed for example subsetting or other binning procedures and what sort of graphical displays and inferential tools will be used!-->


# Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

```{r}
women=triUmp[triUmp$gender == "F",]
```

## Descriptive Results

### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question. Bottom line is to explain the features of the graph that are relevant to conforming or departing to what is expected by your null hypothesis !-->

```{r}
barchartGC(~infected + treatment,data=women)
barchartGC(~infected + treatment,data=women, type="percent")
```

Among women who contracted aids, most of them had undergone treatment. Therefore the treatment was not effective in preventing aids. 

### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r}
table1 <- xtabs(~infected + treatment, data=women)
rowPerc(table1)
colPerc(table1)
```
Among the women infected with aids, 58.82% of them received the treatment.

## Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

The odds ratio show women who took the treat they are 1.54 times more likely to contract aids then those who did not. The Pearson's Chi-squared test set show the P Value is .0000768 so, The fisher Exact P Value is Pearson's is .00006596, so  we reject the null hypothesis because the P value is lower them 0.05. 

# Discussion and Conclusion

<!--Discussion and conclusion here.  If you found a relationship be sure to consider whether the relationship occurs because one of the variavbles causes the other, or whether they perhasps are related for some other reason.  Watch the chapter 6 videos from the GeorgeTown videos collection.!-->

According the the bar chart women who contracted aids, most of them had undergone treatment. Therefore the treatment was not effective in preventing aids.      
The numerical results show women infected with aids, 58.82% of them received the treatment.   
The Pearson's Chi-squared test set show the P Value is .0000768.   
The fisher Exact P Value is Pearson's is .00006596
With this data we reject the null hypothesis because the P value is lower then 0.05. 
